---
title: "Consumption_Community_Ecology"
author: "Emily"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is the Consumption Analysis of the 2022 Community Ecology Lab Data 

<br>

```{r load packages, message = FALSE}
library(tidyverse)
library(dplyr)
library(car)
library(rstatix)
library(knitr)
library(outliers)
```

<br>
## Load Data 

```{r load data, message = FALSE}
raw_data<-read_csv("C:\\github\\Salmonflies\\Data\\communityecologydata_2022.csv", n_max=84)

```
<br>

## Clean Data 

Here, I removed Nas, calculate the proportion eaten and the difference in proportion eaten compared to Cottonwood, and found and labeled outliers

```{r}
clean_LA <- raw_data |>
  select("treatment", "replicate", "leaf type", "leaf area initial", "leaf area final" )|>
  rename("leaftype"="leaf type", "initial_LA"="leaf area initial", "final_LA"="leaf area final")|>
  na.omit()|>
  convert_as_factor(leaftype, treatment)|>
  mutate(prop_eaten=(initial_LA-final_LA)/initial_LA)|>
  group_by(treatment, replicate)|>
  mutate(diffprop=prop_eaten-prop_eaten[leaftype=="Cottonwood"])|>
  ungroup()|>
  group_by(treatment, leaftype)|>
  mutate(IQR = IQR(prop_eaten), 
         o_upper = quantile(prop_eaten, probs=c( .75), na.rm = FALSE)+1.5*IQR,  
            o_lower = quantile(prop_eaten, probs=c( .25), na.rm = FALSE)-1.5*IQR, 
         outlier = if_else( o_lower <= prop_eaten & prop_eaten <= o_upper, F,T))|>
  ungroup()|>
  group_by(treatment, replicate)|>
  mutate(outlier = if_else(any(outlier == 'TRUE'), T, F))|>
  ungroup()


clean_LA |>
  head()|>
  kable()
```




### Check to see counts of each treatment

```{r}
clean_LA|>
  filter(outlier == FALSE)|>
  group_by(treatment)|>
  count() |>
  kable()
```

<br>

# Plotting Proportion Consumed and Difference of Proportion Consumed 

```{r}
clean_LA|>
  ggplot(aes(treatment, prop_eaten, color=leaftype))+
  geom_boxplot()+
  labs(x="Treatment", y="Proportion of Leaf Consumed", color="Leaf Type")+
  theme_bw()

ggsave("Images/Community Ecology Proportion Consumed Boxplot.png")

clean_LA|>
  filter(outlier == FALSE)|>
  ggplot(aes(treatment, prop_eaten, color=leaftype))+
  geom_boxplot()+
  labs(x="Treatment", y="Proportion of Leaf Consumed", color="Leaf Type")+
  theme_bw()

ggsave("Images/Community Ecology Proportion Consumed Boxplot_outliers_rm.png")

```

```{r}
clean_LA|>
  ggplot(aes(treatment, diffprop, color=leaftype))+
  geom_boxplot()+
  labs(x="Treatment", y="Prop. of Other Leaf Consumed - Cottonwood Leaf Consumed", 
       color="Leaf Type")+
  theme_bw()

ggsave("Images/Community Ecology Diff Prop Consumed Boxplot.png")

clean_LA|> 
  filter(outlier == FALSE) |>
  ggplot(aes(treatment, diffprop, color=leaftype))+
  geom_boxplot()+
  labs(x="Treatment", y="Prop. of Other Leaf Consumed - Cottonwood Leaf Consumed", 
       color="Leaf Type")+
  theme_bw()

ggsave("Images/Community Ecology Diff Prop Consumed Boxplot_outliers_rm.png")
```

<br>


## Analysis: Testing for General Differences in Means



###Plot Data 

```{r}
clean_LA|>
  filter(outlier == FALSE)|>
  group_by(treatment, leaftype)|>
  summarize(count=n(), mean=mean(prop_eaten), sd=sd(prop_eaten))|>
  kable()
```



###Check Sample Sizes

```{r}
clean_LA_outlier_rm = clean_LA |>
  filter(outlier == FALSE)
  
  
table(clean_LA_outlier_rm$treatment, clean_LA_outlier_rm$leaftype) |>
  kable()
```

###Check for Normality

I did this since the sample size is small
```{r}
mod=aov(prop_eaten~treatment+leaftype, data=clean_LA_outlier_rm)
qqPlot(mod$residuals)
shapiro.test(mod$residuals) 
```

residuals are normal, p=.5581

<br>

### print model results 
```{r}
summary(mod)

TukeyHSD(mod, conf.level = .95)
```

There is a significant difference in the amount of leaf eaten between treatment and leaf type ( F = 9.761 and p <.0001 under 3 df, and F = 13.009 and p <.00001 under 3 df). 

<br>

## Pairwise T tests using bonferroni correction methods 
```{r}
pairwiseT <-clean_LA |>
  filter(outlier == FALSE) |>
  group_by(treatment)|>
  pairwise_t_test(prop_eaten~leaftype, 
                  paired=TRUE, p.adjust.method = "bonferroni")

pairwiseT|>
  kable()
```

The salmonflies ate significantly more Willow in both treatment C and D (p=.039 under 5 df and p=.017 under 2 df, respectively). The salmonflies ate significantly more Dogwood in treatment D when offered Cottonwood, Willow, and Dogwood (p=.-24 under 5 df), but did not eat significantly more Dogwood in treatment B when only offered Cottonwood or Dogwood (p = .055 under 7df).


