---
title: "Growth Analysis"
author: "Emily"
date: "2024-11-07"
output: github_document
---

## This document is the analysis of the Salmonfly Leaf Experiment 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages, message = FALSE}
library(dplyr)
library(tidyverse)
library(janitor)
```


### PREP DATA

Load the conditioning growth data for both 7 and 25 days of conditioning
```{r load data, message = FALSE}
raw_growth = read_csv("C:\\github\\Salmonflies\\Data\\Allgrowth.csv")


```

<br> 

Cleaning data to make the names better, mutating into correct types, and adding a column to calculate the difference and relative difference of weights. This will be used to later calculate the differences between means within each conditioning period and difference in means between 7 days of conditioning and 25 days of conditioning
```{r clean names, message = FALSE}

clean_growth = raw_growth |>
  clean_names() |>
  rename(tree_species = species, initial_g = initial, final_g = final, 
         conditioning_days = conditioning)|>
  select(conditioning_days, tree_species, initial_g, final_g) |>
  mutate(conditioning_days = as_factor(conditioning_days),
         diff_g= initial_g - final_g, 
         relative_diff = (initial_g - final_g)/initial_g)

seven_growth = clean_growth |>
  filter(conditioning_days == 7)

twentyfive_growth = clean_growth |>
  filter(conditioning_days == 25)
  
```

<br> 

# Analysis


### Average Means 

#### Under 7 Days of Conditioning

```{r average means 7 day}

seven_growth |>
group_by(tree_species) |>
summarize(mean = mean(diff_g))

```

Overall, eating only leaves conditioned for 7 days resulted in a loss of growth. 


#### Under 25 Days of Conditioning

```{r average means 25 day}

twentyfive_growth |>
group_by(tree_species) |>
summarize(mean = mean(diff_g))
```

Individuals lost weight when consuming alder, dogwood, chokecherry, and cottonwood, and gained weight when consuming willow. 


### Testing for differences of growth means between conditioning times 

```{r}
reg3 = lm(diff_g~conditioning_days*tree_species, data= clean_growth)
anova(reg3)


par(mfrow=c(2,2))
plot(reg3)

tapply(X=clean_growth$diff_g, INDEX= clean_growth$tree_species, FUN=sd)

car::leveneTest(aov(diff_g ~ conditioning_days, data=clean_growth)) 

plot(TukeyHSD(aov(reg3)), las=1)
```
We found no significant difference in growth between individuals consuming leaves that had been conditioned for 25 days and leaves that had been conditioned for 7 days (ANOVA F=.6165, P = .65165), though overall conditioning time was also a strong predictor of growth (ANOVA F = 5.1033 P=0.02585).  

<br> 


### ANOVA for relationship between initial weight, tree species, and growth for 7 and 25 days of conditioning

#### Under 7 Days of conditioning

```{r 7 days of conditioning ANOVA growth}

seven_growth_lm = lm(diff_g~initial_g*tree_species, data = seven_growth)

anova(seven_growth_lm)
```

There is no evidence of an interaction between treatment type and initial weight, therefore, treatment type did not affect the growth rate of individuals (F= 1.71, P=0.1633 on 9 and 50 df)

<br> 

#### Under 25 Days of Conditioning

```{r ANOVA for 25 days of conditioning}

twentyfive_growth_lm = lm(diff_g~initial_g*tree_species, data = twentyfive_growth)

anova(twentyfive_growth_lm)
```

We found that initial weight is a strong predictor of salmonfly growth when feeding on 25-day conditioned leaves (ANOVA, P=0.00120, Table 2), regardless of the type of leaf (ANOVA, F= .8556, P=0.497). 

<br>

### ANOVA for relationship between conditioning days and initial mass 

```{r}

conditioning_lm = lm(diff_g~conditioning_days*initial_g, data = clean_growth)

anova(conditioning_lm)

```

There was also a significant interaction between initial weight of the stoneflies and the conditioning period, indicating that initial weight influences the growth of salmonflies differently, depending on conditioning time (ANOVA, F = 9.9786 P=0.00202)

<br> 

### Testing for difference of means between tree species 

#### 7 days of conditioning

```{r diff in means 7 days }

reg1 = lm(diff_g~tree_species, data= seven_growth)

anova(reg1)

par(mfrow=c(2,2))
plot(reg1)

tapply(X=seven_growth$diff_g, INDEX= seven_growth$tree_species, FUN=sd)

car::leveneTest(aov(diff_g ~ tree_species, data=seven_growth)) 

plot(TukeyHSD(aov(reg1)), las=1)
```

Summary statistics of reg1 show that there is no evidence of a difference in growth of salmonflies between leaf species (ANOVA F = .455, P = .768). 

The windows() stuff shows that the data are normal. 

The tapply shows the standard deviations of the data. 

leveneTest shows that the variances are equal (F = .5853, p = .6746)

TukeyHSD shows graphically there are no difference in means. 


#### Under 25 days of Conditioning 


```{r diff in means 25 days }

reg2 = lm(diff_g~tree_species, data= twentyfive_growth)

anova(reg2)

par(mfrow=c(2,2))
plot(reg2)

tapply(X=twentyfive_growth$diff_g, INDEX= twentyfive_growth$tree_species, FUN=sd)

car::leveneTest(aov(diff_g ~ tree_species, data=twentyfive_growth)) 

plot(TukeyHSD(aov(reg2)), las=1)

```
Under 25 days of conditioning, we found no evidence of a difference in average growth between leaf treatment types (ANOVA, F=0.2203, P=0.926)


<br>

### Creating Large Vs Small salmonflies and calculating growth differences between salmonfly size 

#### Under 7 days of conditioning 

```{r}
size_seven_growth = clean_growth |>
  mutate(size = case_when( initial_g > .37 ~ "large", initial_g <.37  ~ "small")) |>
  filter(conditioning_days == 7)


size_7days_lm=lm(diff_g~size, data=size_seven_growth)

anova(size_7days_lm)
```

There was no difference in growth between large (greater than 0.37 g) and small individuals (ANOVA, F= 1.552, p = 0.2179 with 58 df). 


#### Under 25 Days of Conditioning 

```{r}
size_twentyfive_growth = clean_growth |>
  mutate(size = case_when( initial_g > .37 ~ "large", initial_g <.37  ~ "small")) |>
  filter(conditioning_days == 25)


size_25days_lm=lm(diff_g~size, data=size_twentyfive_growth)

anova(size_25days_lm)
```

Size was a strong predictor of salmonfly growth when eating leaves conditioned for 25 days (ANOVA, F = 7.9731, p = .0065 with 58 df)

<br>


## Plots

```{r boxplot relative growth, message = FALSE}
clean_growth |>
  group_by(tree_species)|>
  ggplot(aes(tree_species, relative_diff, color = conditioning_days)) +
  geom_boxplot()+
  labs(x = "Tree Species", y = "Relative Difference of Growth", 
       color = "Leaf Conditioning Time (days)") +
  facet_wrap(~conditioning_days, nrow= 2)+
  theme_bw()

ggsave("Relative Growth Difference in Tree Species.png", path = "Images")
```
<br> 

Scatterplot of Growth vs initial stonefly weight 

```{r, message = FALSE}
clean_growth |>
  group_by(tree_species)|>
  ggplot(aes(initial_g, diff_g, color = tree_species)) +
  geom_point()+
  labs(x = "Initial Weight", y = "Relative Difference of Growth", 
       color = "Tree Species") +
  facet_wrap(~conditioning_days, nrow= 1)+
  theme_bw()

ggsave("Growth Difference vs Initial Weight Separated.png", path = "Images")
```

<br>

Growth vs Initial Weight of stoneflies 

```{r, message = FALSE}
clean_growth |> 
  ggplot(aes(initial_g, diff_g, color = conditioning_days)) +
  geom_point()+
  labs(x = "Initial Weight", y = "Relative Difference of Growth", 
       color = "Conditioning Time (days)") +
  theme_bw()

ggsave("Growth Difference vs Initial Weight.png", path = "Images")

```