---
title: "Growth Analysis"
author: "Emily"
date: "2024-11-07"
output: github_document
---

## This document is the start of attempting again to publish my undergraduate work
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages, message = FALSE}
library(dplyr)
library(tidyverse)
library(janitor)
```


Load the conditioning growth data for both 7 and 25 days of conditioning
```{r load data, message = FALSE}
raw_growth = read_csv("C:\\Users\\emily\\OneDrive\\Documents\\Undergrad\\Undergrad year 4\\Thesis\\Analysis\\Data\\Allgrowth.csv")


```


Cleaning data to make the names better, mutating into correct types, and adding a column to calculate the difference and relative difference of weights. This will be used to later calculate the differences between means within each conditioning period and difference in means between 7 days of conditioning and 25 days of conditioning
```{r clean names, message = FALSE}

clean_growth = raw_growth |>
  clean_names() |>
  rename(tree_species = species, initial_g = initial, final_g = final)|>
  select(conditioning_days, tree_species, initial_g, final_g) |>
  mutate(conditioning_days = as_factor(conditioning_days),
         diff_g= initial_g - final_g, 
         relative_diff = (initial_g - final_g)/initial_g)
  
```
]


```{r boxplot relative growth}
clean_growth |>
  group_by(tree_species)|>
  ggplot(aes(tree_species, relative_diff, color = conditioning_days)) +
  geom_boxplot()+
  labs(x = "Tree Species", y = "Relative Difference of Growth", 
       color = "Leaf Conditioning Time (days)") +
  facet_wrap(~conditioning_days, nrow= 2)+
  theme_bw()
```
<br> 

Scatterplot of Growth vs initial stonefly weight 

```{r}
clean_growth |>
  group_by(tree_species)|>
  ggplot(aes(initial_g, diff_g, color = tree_species)) +
  geom_point()+
  labs(x = "Tree Species", y = "Relative Difference of Growth", 
       color = "Tree Species") +
  facet_wrap(~conditioning_days, nrow= 1)+
  theme_bw()
```

<br>

Growth vs Initial Weight of stoneflies 

```{r}
clean_growth |> 
  ggplot(aes(initial_g, diff_g, color = conditioning_days)) +
  geom_point()+
  labs(x = "Tree Species", y = "Relative Difference of Growth", 
       color = "Conditioning Time (days)") +
  theme_bw()
```


# Calculating ANOVA and LM of difference between means for 7 and 25 days of conditioning on the growth of salmonflies



```{r 7 days of conditioning ANOVA growth}

seven_growth = clean_growth |>
  filter(conditioning_days == 7)

seven_growth_lm = lm(diff_g~initial_g*tree_species, data = seven_growth)

anova(seven_growth_lm)
```
```
Testing for difference of means for between tree species under 25 days of conditioning
```{r ANOVA for 25 days of conditioning}

twentyfive_growth = clean_growth |>
  filter(conditioning_days == 25)

twentyfive_growth_lm = lm(diff_g~initial_g*tree_species, data = twentyfive_growth)

anova(twentyfive_growth_lm)
```


Found that tree species didn't affect growth at 25 days of conditioning, so now testing conditioning days and initial mass 

```{r}

conditioning_lm = lm(diff_g~conditioning_days*initial_g, data = clean_growth)

anova(conditioning_lm)

```

